\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-01-07 archive.tex
hyper parameters and training details

\begin{figure}
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \renewcommand{\inventionColor}{orange}
    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
    }

    \renewcommand{\horizontalSpacing}{3.8cm}
    \renewcommand{\verticalSpacing}{2cm}

@@

\begin{figure}\centering
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \renewcommand{\inventionColor}{orange}

    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$\,=\,$}
    }
    \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
    \renewcommand{\horizontalSpacing}{4.2cm}
    \renewcommand{\verticalSpacing}{2cm}

    \newcommand{\invention}[4]{
      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,draw,rounded corners,align=center] at (#2) {
        \footnotesize #3
      };
      \node[anchor=north,align=center] at ([yshift=3.5pt]#1.south) {#4};
    }
    \newcommand{\vinvention}[4]{
      \invention{#1}{[yshift=-\verticalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\hinvention}[4]{
      \invention{#1}{[xshift=\horizontalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\exampleFunction}[2]{
      \includegraphics[width = 0.5cm]{#1}, \includegraphics[width = 0.5cm]{#2}
    }
    \renewcommand{\task}[4]{
      \node[align=left,anchor=center](t#1) at (#2) {#3};
      \node[align=left](p#1) at ([yshift=-0.75cm]t#1.south) {#4};
    }
    \renewcommand{\nextTask}[4]{
      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

@@

      %% defined routines for creating fragmented syntax trees
      \renewcommand{\syntaxOne}[1]{
        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
          \node(l1) at (0,0) {};
          \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
          \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
          \node(x1) at (0,-1) {\texttt{1}};
          \draw[color=pop3] (l1.south) -- (p1.north);
          \draw[color=pop3] (l1.south) -- (n1.north);
          \draw[color=pop3] (-0.5,-0.45) -- (x1.north);

@@

\subsection{Consolidation-Sleep: Growing a Domain Specific Language}

@@

The second approach is to have $Q$ predict a fixed-dimensional weight
vector, which then biases a fast enumerator~\cite{balog2016deepcoder,ec2}
or sampler~\cite{menon2013machine}.  This approach can enjoy strong
out-of-sample generalization, because it can fall back on enumeration
or sampling when the target program is unlike the training programs.
A main drawback is that the neural net is deliberately handicapped,
and can only send so much information about the target program.

We adopt a middle ground between these two extremes.  Our recognition
model predicts a distribution over primitives in the DSL,
conditioned on the local context in the syntax tree of the
program. When predicting the next node to add to the syntax tree of a program,
the recognition model
conditions on the parent node, as well as
which argument is being generated.
This is a kind of `bigram' model over trees,
where the bigrams
take the form of (parent, child, argument index).
Figure~\ref{bg} diagrams this generative process and Algorithm~\ref{recognitionSample}
specifies a sampling procedure for $Q(\cdot |x)$.
This parameterization
has three main advantages:
(1) it supports fast enumeration
and sampling of programs,
because the
recognition model
only needs to run once for each task;
(2) it allows the recognition model to provide fine-grained
information about the structure of the target program;
and (3)
training this recognition model
causes it to learn to break symmetries in the space of programs,
described next.
\begin{figure}
\centering  \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};

    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
\end{tikzpicture}
\caption{Parameterization of distribution over programs predicted by recognition model.
  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task.}\label{bg}
\end{figure}

@@

\begin{figure}
  \begin{minipage}[c]{0.3\textwidth}
    \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};

    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
\end{tikzpicture}
    \end{minipage}\hfill%
  \begin{tabular}{cll}
     \toprule
&     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
     $\mathcal{L}^{\text{posterior}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{(+ 1 0)}\\
\code{(+ (+ 0 0)}\\\code{\phantom{(+ }(+ 1 0))}\\
\code{(+ 1 1)}
\\
63.0\% right-associative\\37.4\% \code{+0}'s
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{0}\\
\code{(+ (+ (+ 0 0)}\\\code{\phantom{(+ (+} (+ 0 1)) 1)}\\
\code{1}\\
55.8\% right-associative\\31.9\% \code{+0}'s
       \end{tabular}
     \\\\
     $\mathcal{L}^{\text{MAP}}$&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{1}\\
       \code{(+ 1 (+ 1 (+ (+ 1 }\\\code{\phantom{(+ }(+ 1 1)) 1)))}\\
       \code{(+ (+ 1 1) 1)}
\\48.6\% right-associative\\
       0.5\% \code{+0}'s       
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three Samples:}\\
\code{(+ 1 (+ 1 (+ 1}\\\code{\phantom{(+ 1} (+ 1 (+ 1 1)))))}\\
\code{0}\\
\code{(+ 1 (+ 1 (+ 1 1)))}
       \\
       \textbf{97.9\% right-associative}\\
       2.5\% \code{+0}'s       
       \end{tabular}
\\    \bottomrule 
  \end{tabular}
  \caption{\textbf{Left:} Bigram parameterization of distribution over programs predicted by recognition model.
  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task. \textbf{Right:} Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero. \% right-associative calculated by drawing 500 samples from $Q$. $\mathcal{L}^{\text{MAP}}$/Unigram agent incorrectly learns to never generate programs with \code{0}'s, while $\mathcal{L}^{\text{MAP}}$/Bigram agent correctly learns that \code{0} should only be disallowed as an argument of addition. Tasked with building programs from \code{+}, \code{1}, and \code{0}. }\label{symmetry}
\end{figure}

@@

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ec2},
each with 15 input/output examples.  In solving these tasks, the
system composed 16 new DSL components, and discovered multiple
higher-order functions, such as \code{filter} ($f_0$ in
Figure~\ref{initialExampleDSL}).  Each round of memory consolidation
built on components discovered in earlier sleep cycles --- for example
the agent first learns \code{filter}, uses filter to learn to take the
maximum element of a list, then uses that routine to learn a new
component for extracting the $n^{\text{th}}$ largest element of a
list, which it finally uses to learn how to sort a list of numbers
(Figure~\ref{initialExampleDSL}).  This incremental, modular learning
of deep hierarchies of DSL components occurs
because of the alternation between code writing and code refactoring.

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.%%  (e.g., we do not model gravity
%% and discretize the coordinates of the 
%% blocks).
The system starts with the same control flow primitives as with
LOGO graphics,
and learns parametric `options' for building blocks towers (Figure~\ref{tower}),
inferring concepts like arches, staircases, bridges and brick walls.
\begin{figure}
  \begin{tabular}{cc}
    \toprule Example learned DSL components&Samples from DSL\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 5cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{options},
    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', from learned DSL.}\label{tower}
\end{figure}

@@

\begin{figure}
  \begin{tabular}{ccc}
    \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
    \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
  \end{tabular}
  \caption{}\label{learningCurves}
\end{figure}

@@

Our aim with \system is both to show how learning can enable better
program synthesis, and to show how better program synthesis can be
more broadly useful for AI. How could a program induction approach to
AI become broadly useful across a diverse spectrum of AI problems ---
planning, natural language understanding, machine vision, causal
inference, and more?

@@

    \Statex
    \State \textbf{function} sample$'(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
    \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$\Comment{Environment $\mathcal{E}$ starts out as $\varnothing$}
    \State \textbf{Output:} a program whose type unifies with $\tau$
    \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
    \State var $\gets$ an unused variable name
    \State body $\sim$ sample$'(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$\Comment{Recursively sample function body}
    \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
    %\Endif
    \Else\Comment{Build an application to give something w/ type $\tau$}
    \State $\text{primitives} \gets \{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
    $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
    \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
    \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
    \theta_e&\text{ if }e\in \mathcal{D}\\
    \theta_{var}/|\text{variables}|&\text{ if }e\in \mathcal{E}
    \end{cases}$
    \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
    \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
    %   \State unify$(\tau,\beta)$
    \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
    \State $a_k\sim\text{sample}'(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
    \EndFor
    \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
    \EndIf
    \Statex
    \Statex\textbf{where:}
    \State yield$(\tau) = \begin{cases}
      \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
      \tau   &\text{ otherwise.}
    \end{cases}$ \Comment{Final return type of $\tau$}
    \State  args$(\tau) = \begin{cases}
      [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
      []   &\text{ otherwise.}
    \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
  \end{algorithmic}
\end{algorithm}

@@

Figure~\ref{learningFromScratch} shows the full set of tasks and the learned DSL.
\begin{figure*}  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{cc}\toprule
    \normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule

    \begin{tabular}{l}
      \code{[1\, 9]}$\to $\code{2}\\
      \code{[5\, 3\, 8]$\to$3}\\
    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
      \code{[true\, false]}$\to $\code{2}\\
      \code{[false\, false\, false]$\to$3}\\
    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\

\code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
\code{[9\, 8]$\to$[9\, 8\, 0]}\\
\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\

\code{[2\, 1\, 4]$\to$[2\, 1]}\\
%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
\code{[9\, 8]$\to$[9]}\\
\blueCode{$f(\ell) = $($f_0$ ($\lambda$ (z) (empty? (cdr z))) car cdr $\ell$)}\\\\

\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
\blueCode{$f(\ell) = $($f_2$ + $\ell$ 0)}\\\\

>>>>

\begin{figure}
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \newcommand{\inventionColor}{orange}
    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}\,=\,$}}
    }

    \newcommand{\horizontalSpacing}{3.8cm}
    \newcommand{\verticalSpacing}{2cm}

@@

\begin{figure}\centering
  \begin{tikzpicture}
    \renewcommand{\baselinestretch}{0.5}
    \newcommand{\inventionColor}{orange}

    \newcommand{\name}[2]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$($#2$)\,=\,$}
    }
    \newcommand{\noArguments}[1]{
      \mcode{{\color{\inventionColor}$f_{#1}$}$\,=\,$}
    }
    \newcommand{\use}[1]{{\color{\inventionColor}$f_{#1}$}}
    \newcommand{\horizontalSpacing}{4.2cm}
    \newcommand{\verticalSpacing}{2cm}

    \newcommand{\invention}[4]{
      \node(#1)[inner sep=1,anchor=south,text width=2.2cm,draw,rounded corners,align=center] at (#2) {
        \footnotesize #3
      };
      \node[anchor=north,align=center] at ([yshift=3.5pt]#1.south) {#4};
    }
    \newcommand{\vinvention}[4]{
      \invention{#1}{[yshift=-\verticalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\hinvention}[4]{
      \invention{#1}{[xshift=\horizontalSpacing]#2.south}{#3}{#4}
    }
    \newcommand{\exampleFunction}[2]{
      \includegraphics[width = 0.5cm]{#1}, \includegraphics[width = 0.5cm]{#2}
    }
    \newcommand{\task}[4]{
      \node[align=left,anchor=center](t#1) at (#2) {#3};
      \node[align=left](p#1) at ([yshift=-0.75cm]t#1.south) {#4};
    }
    \newcommand{\nextTask}[4]{
      \task{#1}{[xshift=\horizontalSpacing]#2.center}{#3}{#4}
    }

@@

      %% defined routines for creating fragmented syntax trees
      \newcommand{\syntaxOne}[1]{
        \begin{tikzpicture}[scale=#1,line width=0.35mm]          
          \node(l1) at (0,0) {};
          \node[color=pop3](p1) at (-1,-1) {\texttt{+}};
          \node[color=pop3](n1) at (0.7,-0.9) {\texttt{1}};
          \node(x1) at (0,-1) {\texttt{1}};
          \draw[color=pop3] (l1.south) -- (p1.north);
          \draw[color=pop3] (l1.south) -- (n1.north);
          \draw[color=pop3] (-0.5,-0.45) -- (x1.north);

@@

\subsection{Consolidation-Sleep: Growing a Domain Specific Language}\label{consolidationSection}

@@

The second approach is to have $Q$ predict a fixed-dimensional weight
vector, which then biases a fast enumerator~\cite{balog2016deepcoder,ecc}
or sampler~\cite{menon2013machine}.  This approach can enjoy strong
out-of-sample generalization, because it can fall back on enumeration
or sampling when the target program is unlike the training programs.
A main drawback is that the neural net is deliberately handicapped,
and can only send so much information about the target program.

We adopt a middle ground between these two extremes.  Our recognition
model predicts a distribution over primitives in the DSL,
conditioned on the local context in the syntax tree of the
program. When predicting the next node to add to the syntax tree of a program,
the recognition model
conditions on the parent node, as well as
which argument is being generated.
This is a kind of `bigram' model over trees,
where the bigrams
take the form of (parent, child, argument index).
Figure~\ref{symmetry} (left) diagrams this generative process and Algorithm~\ref{recognitionSample}
specifies a sampling procedure for $Q(\cdot |x)$.
This parameterization
has three main advantages:
(1) it supports fast enumeration
and sampling of programs,
because the
recognition model
only needs to run once for each task;
(2) it allows the recognition model to provide fine-grained
information about the structure of the target program;
and (3)
training this recognition model
causes it to learn to break symmetries in the space of programs,
described next.

@@

\begin{figure}
  \begin{minipage}[c]{0.3\textwidth}
    \begin{tikzpicture}[scale=1]%[every node/.style={inner sep=0,outer sep=0}]
    \node(l)[draw] at (0,0) {\texttt{$\lambda$ (a)}};
    \node(k)[draw] at ([yshift=-50]l.south) {\texttt{+}};
    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\texttt{9}};
    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\texttt{*}};
    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\texttt{a}};
    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\texttt{a}};

    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$Q_{\text{start},\texttt{+},1}(x)$}(k.north);
    \draw[->] (k.south)--node[sloped, above]{$Q_{\texttt{+},\texttt{9},1}(x)$}(o.north);
    \draw[->] (k.south)--node[sloped, above]{$Q_{\texttt{+},\texttt{*},2}(x)$}(m.north);
    \draw[->] (m.south)--node[sloped, above]{$Q_{\texttt{*},\text{var},2}(x)$}(x1.north);
    \draw[->] (m.south)--node[sloped, above]{$Q_{\texttt{*},\text{var},1}(x)$}(x2.north);
\end{tikzpicture}
    \end{minipage}\hfill%
  \begin{tabular}{cll}
     \toprule
&     \multicolumn{1}{c}{Unigram}&\multicolumn{1}{c}{Bigram} \\\midrule
     \rotatebox[origin=c]{90}{$\mathcal{L}^{\text{posterior}}$}&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{(+ 1 0)}\\
\code{(+ (+ 0 0)}\\\code{\phantom{(+ }(+ 1 0))}\\
\code{(+ 1 1)}
\\
63.0\% right-associative\\37.4\% \code{+0}'s
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{0}\\
\code{(+ (+ (+ 0 0)}\\\code{\phantom{(+ (+} (+ 0 1)) 1)}\\
\code{1}\\
55.8\% right-associative\\31.9\% \code{+0}'s
       \end{tabular}
     \\\\
     \rotatebox[origin=c]{90}{$\mathcal{L}^{\text{MAP}}$}&
     \begin{tabular}{l}
       \emph{Three samples:}\\
       \code{1}\\
       \code{(+ 1 (+ 1 (+ (+ 1 }\\\code{\phantom{(+ }(+ 1 1)) 1)))}\\
       \code{(+ (+ 1 1) 1)}
\\48.6\% right-associative\\
       0.5\% \code{+0}'s       
       \end{tabular}
     &
     \begin{tabular}{l}
       \emph{Three Samples:}\\
\code{(+ 1 (+ 1 (+ 1}\\\code{\phantom{(+ 1} (+ 1 (+ 1 1)))))}\\
\code{0}\\
\code{(+ 1 (+ 1 (+ 1 1)))}
       \\
       \textbf{97.9\% right-associative}\\
       2.5\% \code{+0}'s       
       \end{tabular}
\\    \bottomrule 
  \end{tabular}
  \caption{\textbf{Left:} Bigram parameterization of distribution over programs predicted by recognition model.
  Here the program (syntax tree shown above) is \texttt{($\lambda$ (a) (+ 9 (* a a )))}.
Each conditional distribution predicted by the recognition model is written $Q_{\text{parent},\text{child},\text{argument index}}(x)$, where $x$ is a task. \textbf{Right:} Agent learns to break symmetries in program space only when using both bigram parameterization and $\mathcal{L}^{\text{MAP}}$ objective, associating addition to the right and avoiding adding zero. \% right-associative calculated by drawing 500 samples from $Q$. $\mathcal{L}^{\text{MAP}}$/Unigram agent incorrectly learns to never generate programs with \code{0}'s, while $\mathcal{L}^{\text{MAP}}$/Bigram agent correctly learns that \code{0} should only be disallowed as an argument of addition. Tasked with building programs from \code{+}, \code{1}, and \code{0}. }\label{symmetry}
\end{figure}

@@

\subsubsection{List Processing}\label{listSection}
We took 218 list manipulation tasks from our previous work~\cite{ecc},
each with 15 input/output examples.  In solving these tasks, the
system composed 16 new DSL components, and discovered multiple
higher-order functions, such as \code{filter} ($f_0$ in
Figure~\ref{initialExampleDSL}).  Each round of memory consolidation
built on components discovered in earlier sleep cycles --- for example
the agent first learns \code{filter}, uses filter to learn to take the
maximum element of a list, then uses that routine to learn a new
component for extracting the $n^{\text{th}}$ largest element of a
list, which it finally uses to learn how to sort a list of numbers
(Figure~\ref{initialExampleDSL}).  This incremental, modular learning
of deep hierarchies of DSL components occurs
because of the alternation between code writing and code refactoring.

@@

Inspired by the classic AI `copy task' --- where an agent must look at
an image of a tower made of toy blocks and re-create the
tower~\cite{towerCopy} --- we give \system 112 tower `copy tasks'
(Figure~\ref{towerTasks}).  Here the agent observes both an image of a
tower and the locations of each of its blocks, and must write a
program that plans how a simulated hand would build the tower.
These towers are built from Lego-style blocks
that snap together on a discrete grid.%%  (e.g., we do not model gravity
%% and discretize the coordinates of the 
%% blocks).
The system starts with the same control flow primitives as with
LOGO graphics,
and learns parametric `options' for building blocks towers (Figure~\ref{tower}),
inferring concepts like arches, staircases, bridges and brick walls.
\begin{figure}
  \begin{tabular}{cc}
    \toprule Example learned DSL components&Samples from DSL\\\midrule
    \begin{tabular}{rl}
      Brickwall&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bricks.png}
      \end{tabular}\\
      Bridge&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_bridge.png}
      \end{tabular}\\
      Staircase&\begin{tabular}{l}
        \includegraphics[width = 5cm]{figures/tower/tower_dsl_staircase.png}
        \end{tabular}
    \end{tabular}&
    \begin{tabular}{l}
      \includegraphics[width = 5cm]{figures/tower/dreams.png}
      \end{tabular}
    \\\bottomrule 
    \end{tabular}
  \caption{\textbf{Left}: Three (out of 19) learned DSL components for building towers out of Lego-style blocks. These components act like parametric options~\cite{stolle2002learning},
    giving higher-level building blocks that the agent can use to plan. \textbf{Right}: 16 random samples, or `dreams', from learned DSL.}\label{tower}
\end{figure}

@@

\begin{figure}
  \begin{tabular}{ccc}
    \includegraphics[width = 5cm]{figures/logoLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/textLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/listLearningCurve_batch.png}\\\\
    \includegraphics[width = 5cm]{figures/towerLearningCurve_batch.png}&
    \includegraphics[width = 5cm]{figures/rationalLearningCurve_batch.png}&
  \end{tabular}
  \caption{Test set performance across wake/sleep iterations. An Epoch is a full pass over the set of training tasks. Orange: \systemEnding. Teal: No recognition model. Purple: No DSL learning. Solid lines: \% tasks solved within 10 minute timeout. Dashed lines: average search time. Shaded region is $\pm 1$ stddev search time.}\label{learningCurves}
\end{figure}

@@

Our aim with \system is both to show how learning can enable better
program synthesis, and to show how better program synthesis can be
more broadly useful for AI.
Many challenges remain.

How could a program induction approach to
AI become broadly useful across a diverse spectrum of AI problems ---
planning, natural language understanding, machine vision, causal
inference, and more?

@@

    \Statex
    \State \textbf{function} sample$'(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
    \State {\bfseries Input:} DSL $(\mathcal{D},\theta)$, environment $\mathcal{E}$, type $\tau$\Comment{Environment $\mathcal{E}$ starts out as $\varnothing$}
    \State \textbf{Output:} a program whose type unifies with $\tau$
    \If{$\tau = \alpha\to\beta$}\Comment{Function type --- start with a lambda}
    \State var $\gets$ an unused variable name
    \State body $\sim$ sample$'(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$\Comment{Recursively sample function body}
    \State \textbf{return} \code{(lambda (}var\code{) }body\code{)}
    %\Endif
    \Else\Comment{Build an application to give something w/ type $\tau$}
    \State $\text{primitives} \gets \{p | p: \tau' \in \mathcal{D}\cup\mathcal{E}$
    $\text{if }\tau\text{ can unify with yield}(\tau') \}$\Comment{Everything in scope w/ type $\tau$}
    \State $\text{variables}\gets\left\{p\;|\;p\in \text{primitives}\text{ and }p\text{ a variable} \right\}$
    \State Draw $e\sim \text{primitives}$, w.p. $\propto\begin{cases}
    \theta_e&\text{ if }e\in \mathcal{D}\\
    \theta_{var}/|\text{variables}|&\text{ if }e\in \mathcal{E}
    \end{cases}$
    \State Unify $\tau$ with yield$(\tau')$.\Comment{Ensure well-typed program}
    \State $\left\{\alpha_k \right\}_{k = 1}^K\gets\text{args}(\tau')$ 
    %   \State unify$(\tau,\beta)$
    \For{$k=1$ {\bfseries to} $K$}\Comment{Recursively sample arguments}
    \State $a_k\sim\text{sample}'(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
    \EndFor
    \State \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
    \EndIf
    \Statex
    \Statex\textbf{where:}
    \State yield$(\tau) = \begin{cases}
      \text{yield}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
      \tau   &\text{ otherwise.}
    \end{cases}$ \Comment{Final return type of $\tau$}
    \State  args$(\tau) = \begin{cases}
      [\alpha] + \text{args}(\beta)   &\text{ if }\tau = \alpha\to \beta\\
      []   &\text{ otherwise.}
    \end{cases}$\Comment{Types of arguments needed to get something w/ type $\tau$}
  \end{algorithmic}
\end{algorithm}

@@

Figure~\ref{learningFromScratch} shows the full set of tasks and the learned DSL.
\begin{figure}  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{ll}\toprule
    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&%&\normalsize \popp{DSL}\\\midrule

    \begin{tabular}{ll}      
      \begin{tabular}{l}
        \code{[1\, 9]}$\to $\code{2}\\
        \code{[5\, 3\, 8]$\to$3}\\
        \blueCode{$f(\ell) = $($f_5$ $\ell$)}
      \end{tabular}&
      \begin{tabular}{l}
        \code{[[2 1]\, []]}$\to $\code{[1 0]}\\
        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
        \blueCode{$f(\ell) = $($f_2$ $f_4$ $\ell$)}
      \end{tabular}\\\\

      \begin{tabular}{l}
        \code{[0 1 1 0 0]$\to$[1 1]}\\
        \code{[9 0 8]$\to$[9\, 8]}\\
        \blueCode{$f(\ell) = $($f_3$ (eq? 0) $\ell$)}        
      \end{tabular}&
      \begin{tabular}{l}
        \code{[1 -1 0 2]$\to$[1 2]}\\
        \code{[9 -5 5 0 8]$\to$[9 5 8]}\\
        \blueCode{$f(\ell) = $($f_3$ (gt? 1) $\ell$)}        
      \end{tabular}\\\\

      \begin{tabular}{l}
        \code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
        \code{[9\, 8]$\to$[9\, 8\, 0]}\\
        \blueCode{$f(\ell) = $($f_0$ cons $\ell$ (cons 0 nil))}
      \end{tabular}&

      \begin{tabular}{l}
        \code{[2\, 1\, 4]$\to$[2\, 1]}\\
        \code{[9\, 8]$\to$[9]}\\
        \blueCode{$f(\ell) = $($f_6$ cdr ($\lambda$ (z) (empty? (cdr z))) $\ell$)}
      \end{tabular}\\\\

      \begin{tabular}{l}
        \code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
        \code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
        \blueCode{$f(\ell) = $($f_0$ + $\ell$ 0)}
      \end{tabular}&

      \begin{tabular}{l}
        \code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
        \code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
        \blueCode{$f(\ell) = $($f_2$ ($\lambda$ (x) (+ x x)) $\ell$)}
      \end{tabular}\\\\
